Imputation of Missing Values in the Melbourne CBD Pedestrian Dataset

### 2017-08-20 thesis/missing_impute.Rmd
updated with knitted html

---
title: "Imputation of Missing Values in the Melbourne CBD Pedestrian Dataset"
output:
  html_document: default
  html_notebook: default
---
```{r, include=FALSE}
library(plyr)
library(tidyverse)
library(lubridate)
library(magrittr)
library(viridis)
library(gridExtra)
set.seed(12345)
```

@@

```
Using the coefficients saved for each location, we can easily fit the test data.  
With the function `model.matrix()`, `R` will generate a matrix, $X$, with the factor levels converted into dummy variables for use with the GLM. Then, we can obtain our estimates of $\text{Hourly_Counts}_{ij}$, $\widehat{\text{Hourly_Counts}_i} = \exp \left( X \hat{\beta_j} \right)$ where $\hat{\beta_j}$ is a column vector of the GLM estimated coefficients at the $j^{th}$ sensor location.

# Goodness of Fit
### Visually
```{r, fig.align= 'center', fig.height=24, fig.width=10, fig.retina=4, cache=TRUE, echo=FALSE}
ggplot(filter(test_fitted, Date > "2017-07-23")) +
  geom_line(aes(x = Date_Time, y = Hourly_Counts)) +
  geom_line(aes(x = Date_Time, y = Fitted),
            colour = "violetred2", alpha = 0.75) +
                         facet_wrap(~ Sensor, scales = "free_y", ncol = 4) + theme_minimal() + 
                    theme(axis.text.x = element_text(size = 3),
                          strip.text = element_text(size = 10),
                          axis.text.y = element_text(size = 0)) +
                    xlab("Date") + ylab("Hourly Counts")
```
Looking at plots of actuals (black) against fitted (red) of data from 24 July 2017 to 6 August 2017, we see that some sensors appear to be not operational. We see that there is a mix of very good fitting models and not so good ones. An example of of a reasonably good fitting model is the _Collins Place (North)_ sensor location:
```{r, fig.align= 'center', fig.retina = 4, cache=TRUE}
ggplot(test_fitted %>%
         filter(Date > "2017-07-23",
                Sensor == "Collins Place (North)")) +
      geom_line(aes(x = Date_Time, y = Hourly_Counts)) +
      geom_line(aes(x = Date_Time, y = Fitted), colour = 'violetred2', alpha = 0.7) +
      theme_minimal()
```

@@

```{r, echo=FALSE, fig.align='center', fig.retina=4, cache=TRUE}
ggplot(test_fitted %>%
         filter(Date > "2017-07-23",
                Sensor == "Grattan St-Swanston St (West)")) +
      geom_line(aes(x = Date_Time, y = Hourly_Counts)) +
      geom_line(aes(x = Date_Time, y = Fitted), colour = 'violetred2', alpha = 0.7) +
      theme_minimal()
```

@@

```{r, fig.align= 'center', fig.retina=3, echo=FALSE, cache=TRUE}
ggplot(filter(test_fitted, Sensor_ID == "Grattan St-Swanston St (West)")) +
  geom_path(aes(x = Time, y = Hourly_Counts, group = Date), alpha = 0.1) +
  geom_path(aes(x = Time, y = Fitted, group = Date), colour = "violetred2", alpha = 0.1) +
  facet_wrap(~ DayType) +
      theme_minimal()
```

@@

```{r, fig.align= 'center', fig.retina=3, echo=FALSE, cache=TRUE}
ggplot(filter(ped_data, Sensor_ID == "Grattan St-Swanston St (West)")) +
  geom_path(aes(x = Time, y = Hourly_Counts, group = Date, colour = Month), alpha = 0.1) +
  facet_wrap(~ DayType) +
      theme_minimal()

@@

However, we see that the variance within a month is quite low comparatively. It is worth noting that the data available at this location is very poor.

```{r}

```

>>>>

---
title: "Imputation of Missing Values in the Melbourne CBD Pedestrian Dataset"
output:
  html_document:
    theme: united
  html_notebook: default
  pdf_document: default
---
```{r, include=FALSE}
library(plyr)
library(tidyverse)
library(lubridate)
library(magrittr)
library(viridis)
library(gridExtra)
set.seed(12345)
```

@@

```
Using the coefficients saved for each location, we can easily fit the test data.  
With the function `model.matrix()`, `R` will generate a matrix, $X_j = \{ X_{1j}, X_{2j}, \dots , X_{Tj} \}$, with the factor levels converted into dummy variables for use with the GLM. Then, we can obtain our estimates of $\text{Hourly_Counts}_{j}$, $\widehat{\text{Hourly} \_ \text{Counts}}_{j} = \exp \left( X_j\hat{\beta_j} \right)$ where $\hat{\beta_j}$ is a column vector of the GLM estimated coefficients at the $j^{th}$ sensor location.

# Goodness of Fit
### Visually
```{r, fig.align= 'center', fig.height=24, fig.width=10, fig.retina=1, cache=TRUE, echo=FALSE}
ggplot(filter(test_fitted, Date > "2017-07-23")) +
  geom_line(aes(x = Date_Time, y = Hourly_Counts)) +
  geom_line(aes(x = Date_Time, y = Fitted),
            colour = "violetred2", alpha = 0.75) +
                         facet_wrap(~ Sensor, scales = "free_y", ncol = 4) + theme_minimal() + 
                    theme(axis.text.x = element_text(size = 3),
                          strip.text = element_text(size = 10),
                          axis.text.y = element_text(size = 0)) +
                    xlab("Date") + ylab("Hourly Counts")
```
Looking at plots of actuals (black) against fitted (red) of data from 24 July 2017 to 6 August 2017, we see that some sensors appear to be not operational. We see that there is a mix of very good fitting models and not so good ones. An example of of a reasonably good fitting model is the _Collins Place (North)_ sensor location:
```{r, fig.align= 'center', fig.retina = 1, cache=TRUE}
ggplot(test_fitted %>%
         filter(Date > "2017-07-23",
                Sensor == "Collins Place (North)")) +
      geom_line(aes(x = Date_Time, y = Hourly_Counts)) +
      geom_line(aes(x = Date_Time, y = Fitted), colour = 'violetred2', alpha = 0.7) +
      theme_minimal()
```

@@

```{r, echo=FALSE, fig.align='center', fig.retina=1, cache=TRUE}
ggplot(test_fitted %>%
         filter(Date > "2017-07-23",
                Sensor == "Grattan St-Swanston St (West)")) +
      geom_line(aes(x = Date_Time, y = Hourly_Counts)) +
      geom_line(aes(x = Date_Time, y = Fitted), colour = 'violetred2', alpha = 0.7) +
      theme_minimal()
```

@@

```{r, fig.align= 'center', fig.retina=1, echo=FALSE, cache=TRUE, warning=F}
ggplot(filter(test_fitted, Sensor_ID == "Grattan St-Swanston St (West)")) +
  geom_path(aes(x = Time, y = Hourly_Counts, group = Date), alpha = 0.1) +
  geom_path(aes(x = Time, y = Fitted, group = Date), colour = "violetred2", alpha = 0.1) +
  facet_wrap(~ DayType) +
      theme_minimal()
```

@@

```{r, fig.align= 'center', fig.retina=1, echo=FALSE, cache=TRUE, warning = F}
ggplot(filter(ped_data, Sensor_ID == "Grattan St-Swanston St (West)")) +
  geom_path(aes(x = Time, y = Hourly_Counts, group = Date, colour = Month), alpha = 0.1) +
  facet_wrap(~ DayType) +
      theme_minimal()

@@

However, we see that the variance within a month is quite low comparatively. It is worth noting that the data available at this location is very poor.  
Plotting all available observations, we see we only have data starting from `r head(na.omit(ped_data[ped_data$Sensor_ID == "Grattan St-Swanston St (West)", ])$Date, 1)`.
